---
title: "Project"
author: "Teshale"
date: "`r Sys.Date()`"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

# Introduction 

* This project aims to analyze board game data from BoardGameGeek to uncover insights regarding game popularity, player preferences, and trends. By understanding these patterns, we can identify what makes a board game successful and how preferences have evolved over the years. 
* As the popularity of board games grows, there is a need to understand the factors that contribute to their success. This analysis seeks to uncover the patterns and trends in board game data, focusing on ratings, player count, and categories. 
* We will use data from BoardGameGeek, including game ratings, categories, mechanics, and publication years. The analysis will involve data cleaning, exploratory data analysis (EDA), and visualization techniques to derive meaningful insights. 
* The findings from this analysis can help game designers, publishers, and enthusiasts understand what makes a board game popular and how player preferences have changed over time.

# Data 

```{r}
library(tidyverse) 
#game_data <- read_csv("BGGdata.csv", show_col_types = FALSE) 
game_data <- read_csv("bgg_db_2017_04.csv/bgg_db_2017_04.csv")
board_game <- read.csv("board_game_popularity_over_years.csv") 
#artists_reduced <- read.csv("project/artists_reduced.csv/artists_reduced.csv") 
glimpse(game_data) 
glimpse(board_game) 
#glimpse(artists_reduced)

```

# the populartiy of the game over a year

```{r}
board_game %>% ggplot(aes(x = yearpublished, y = count)) +
  geom_line() +
  labs(title = "Popularity of Games Over the Years", x = "Year Published", y = "Number of Games")

```
# How many games are in the dataset??

```{r}
# Count the number of rows (games) in the dataset
num_games <- nrow(game_data)
print(paste("Number of games:", num_games))

```


# Calculate the average minimum and maximum number of players

```{r}
average_minplayers <- mean(game_data$minplayers, na.rm = TRUE)
average_maxplayers <- mean(game_data$maxplayers, na.rm = TRUE)
print(paste("Average minimum number of players:", average_minplayers))
print(paste("Average maximum number of players:", average_maxplayers))

```
# highest-rated board games

```{r}

top_rated_games <- game_data %>%
  arrange(desc(avg_rating)) %>%
  select(names, avg_rating) %>%
  head(10)

top_rated_games

```
# Plot game ratings by player count

```{r}
ggplot(game_data, aes(x = factor(min_players), y = avg_rating)) +
  geom_boxplot() +
  labs(title = "Game Ratings by Player Count", x = "Minimum Players", y = "Average Rating")


```
# # Scatter plot of player count vs. ratings
```{r}

ggplot(game_data, aes(x = min_players, y = avg_rating)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Player Count vs. Ratings", x = "Minimum Players", y = "Average Rating")

```

# Identify trends in game publication over the years.

```{r}
ggplot(board_game, aes(x = yearpublished, y = count)) +
  geom_col(fill = "blue", color = "black") +
  labs(title = "Number of Games Published Over Time", x = "Year", y = "Number of Games")

```
# Calculate key statistics for game ratings, number of votes, and playing times.

```{r}
summary(game_data)

```
# Determine the most 10 popular game categories?

```{r}

game_data$category <- as.character(game_data$category)
count_categories <- function(data) {
  split_categories <- strsplit(data$category, ", ")
  categories_list <- unlist(split_categories)
  categories_df <- data.frame(category = categories_list)
  
  # Count the occurrences of each category
  top_categories <- categories_df %>%
    count(category, sort = TRUE) %>%
    head(10)
  
  return(top_categories)
}
top_categories <- count_categories(game_data)

print(top_categories)



```

# Create bar charts to show the popularity of different game categories?

```{r}
ggplot(top_categories, aes(x = reorder(category, n), y = n)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Top 10 Game Categories", x = "Category", y = "Count") +
  coord_flip()
```


# Visualize the distribution of game ratings using histograms.

```{r}
ggplot(game_data, aes(x = avg_rating)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  labs(title = "Distribution of Game Ratings", x = "Average Rating", y = "Frequency")

```
```{r}


```


