---
title: "Project"
author: "Mekonnen Teshale"
date: "`r Sys.Date()`"
output: slidy_presentation
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

# Introduction:  

* This project aims to analyze board game data from BoardGameGeek to uncover insights regarding game popularity, player preferences, and trends. By understanding these patterns, we can identify what makes a board game successful and how preferences have evolved over the years. 
* As the popularity of board games grows, there is a need to understand the factors that contribute to their success. This analysis seeks to uncover the patterns and trends in board game data, focusing on ratings, player count, and categories. 
* We will use data from BoardGameGeek, including game ratings, categories, mechanics, and publication years. The analysis will involve data cleaning, exploratory data analysis (EDA), and visualization techniques to derive meaningful insights. 
* The findings from this analysis can help game designers, publishers, and enthusiasts understand what makes a board game popular and how player preferences have changed over time.

# Data Preparation:

* Source: The data was obtained from Kaggle. 
* (https://www.kaggle.com/code/devisangeetha/insights-geek-board-game/input).
* Data Explanation: 
* The data is collected by BoardGameGeek, a popular online database for board games.
* To provide comprehensive information about board games, including ratings, categories, and mechanics.
* The dataset includes data up to April 2017.
* The data may contain missing values and entries that need to be cleaned for accurate analysis.

# Data import

```{r}
library(tidyverse) 
#game_data <- read_csv("BGGdata.csv", show_col_types = FALSE) 
game_data <- read_csv("bgg_db_2017_04.csv/bgg_db_2017_04.csv")
board_game <- read.csv("board_game_popularity_over_years.csv") 
#artists_reduced <- read.csv("project/artists_reduced.csv/artists_reduced.csv") 
glimpse(game_data) 
glimpse(board_game) 
#glimpse(artists_reduced)

```



# Exploratory Data Analysis:

* key statistics for game ratings, number of votes, and playing times.

```{r}
summary(game_data)
```

# Distribution of Ratings: 
* Understand the distribution of average ratings to identify common rating patterns.

1, the nunber of games in the dataset.

```{r echo = F, message=T}
# Count the number of rows (games) in the dataset
num_games <- nrow(game_data)
print(paste("Number of games:", num_games))

```
2, the average minimum and maximum number of players.

```{r echo = F, message=T}
average_minplayers <- mean(game_data$min_players, na.rm = TRUE)
average_maxplayers <- mean(game_data$max_players, na.rm = TRUE)
print(paste("Average minimum number of players:", average_minplayers))
print(paste("Average maximum number of players:", average_maxplayers))


```
3, Top 10 highest-rated board games.

```{r echo = F, message=T}

top_rated_games <- game_data %>%
  arrange(desc(avg_rating)) %>%
  select(names, avg_rating) %>%
  head(10)

top_rated_games

```

# 4, Plotting game ratings by player count.

```{r echo = F, message=T}
ggplot(game_data, aes(x = factor(min_players), y = avg_rating)) +
  geom_boxplot() +
  labs(title = "Game Ratings by Player Count", x = "Minimum Players", y = "Average Rating")


```

* Ratings tend to vary with the number of players, showing interesting trends in player preferences.



# 5, Scatter plot of player count vs. ratings.

```{r echo = F, message=T}

ggplot(game_data, aes(x = min_players, y = avg_rating)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Player Count vs. Ratings", x = "Minimum Players", y = "Average Rating")

```

* The scatter plot and trend line reveal the correlation between player count and average rating.



# Publication Trends: 
* Examine how the number of games published each year has changed over time.

1, the published of the game over a year

```{r echo = F, message=T}
board_game %>% ggplot(aes(x = yearpublished, y = count)) +
  geom_line() +
  labs(title = "Popularity of Games Over the Years", x = "Year Published", y = "Number of Games")
```

* There is a clear trend in the number of games published over the years.



# 2, Identify trends in game publication over the years.

```{r echo = F, message=T}
ggplot(board_game, aes(x = yearpublished, y = count)) +
  geom_col(fill = "blue", color = "black") +
  labs(title = "Number of Games Published Over Time", x = "Year", y = "Number of Games")

```

* The histogram shows trends and peaks in game publications over time.




# Category Breakdown: 
* Analyze the proportion of different game categories to see which are the most popular.

1, the most 10 popular game categories?

```{r echo = F, message=T}

game_data$category <- as.character(game_data$category)
count_categories <- function(data) {
  split_categories <- strsplit(data$category, ", ")
  categories_list <- unlist(split_categories)
  categories_df <- data.frame(category = categories_list)
  
  # Count the occurrences of each category
  top_categories <- categories_df %>%
    count(category, sort = TRUE) %>%
    head(10)
  
  return(top_categories)
}
top_categories <- count_categories(game_data)

print(top_categories)



```

# 2, bar charts to show the popularity of different game categories?

```{r echo = F, message=T}
ggplot(top_categories, aes(x = reorder(category, n), y = n)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Top 10 Game Categories", x = "Category", y = "Count") +
  coord_flip()
```

* Bar charts provide a clear visual representation of the most popular game categories.



# 3, Visualize the distribution of game ratings using histograms.

```{r echo = F, message=T}
ggplot(game_data, aes(x = avg_rating)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  labs(title = "Distribution of Game Ratings", x = "Average Rating", y = "Frequency")

```

* The histogram highlights the distribution of game ratings across the dataset.



# Summary: 

* This project aimed to analyze board game data to uncover insights into game popularity, player preferences, and trends
* Using data from BoardGameGeek, we performed data cleaning, exploratory data analysis, and visualization to derive meaningful insights.

  # Key Insights:
   
* The popularity of board games has increased over the years.
* The highest-rated games have distinct characteristics.
* Certain categories and mechanics are more popular than others.
* These findings can help game designers, publishers, and enthusiasts understand what makes a board game popular and how player preferences have changed over time.
* The analysis is based on data up to April 2017, and newer trends may not be reflected. Future analysis could include more recent data.



# Citations

* Kaggle Dataset:
* https://www.kaggle.com/code/devisangeetha/insights-geek-board-game/input


